\chapter*{Acknowledgments}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

The work in this thesis was only possible because of all the helpful and guiding members of the UMass Lowell engineering education community. Many people have helped me not just in research effort, but in introducing me to new thoughts and ideas that became foundational to this project.

I would like to first thank Dr. Michelle Scribner-MacLean for being so inclusive towards me. She involved me in her work, where I learned about many of the foundational works that appear in this document. It was not just about getting sources for the literature review, it was about having an experienced colleague.  She taught me that everything is always a learning process, for students, teachers, and researchers alike.

I give my most sincere thanks to Dr. Fred Martin. He was patient and open-minded, yet always loyal to the science and process of development. I would like to thank Dr. Martin for years he has already invested in me, the care he has taken to see me succeed, and the all the opportunities he has afforded me. 

I thank Howard Sticklor, who set up the program upon which this research is based. I very much appreciate his support of both my work and that of our student participants.

I extend thanks to Dr. Sarah Kuhn and Michael Penta, with both of whom I have shared many exciting conversations on education methods and research.  Your ideas are woven into this thesis just as much as my own.

I thank my parents, Karen and Barry, who have been supportive and encouraging in all my endeavors, no matter how daunting. I believe that their continued reinforcement (and often reality checking)  has been critical in all of my successes. They sparked my interest in learning and teaching early on in my life. Their stories are inspirational to me, and I try every day to be as hard working as they are.

I would lastly like to thank my closest friends who have supported me through every phase of this process. Nick McKinnon, Mary Angeleri, and especially Stacy Kadesch, I thank you.

This material is based upon work supported by the National Science Foundation under Grants No. DRL-0624669 and No. DGE-0841392.\footnote{Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the National Science Foundation.}